---
title: "RepData_PeerAssessment1"
author: "Corbin Charpentier"
date: "4/20/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
```

## Loading and preprocessing the data

```{r data}
data <- read.csv("activity.csv")
completeData <- data[complete.cases(data), ]
```

## What is mean total number of steps taken per day?
```{r daily steps, message=F, warning=F}
dailySteps <- group_by(completeData, date) %>% summarize(steps=sum(steps))
qplot(dailySteps$steps, xlab="Number of Steps", ylab="Frequency", binwidth=1000)
```

### Mean steps per day
```{r mean steps}
mean <- mean(dailySteps$steps, na.rm=T)
median <- median(dailySteps$steps, na.rm=T)
```
The average number of daily steps is `r mean` and the median is `r median`. 

## What is the average daily activity pattern?
### Graph daily interval against average steps for each interval
```{r interval steps}
avgPerInterval <- group_by(completeData, interval) %>% summarize(steps=mean(steps))
with(avgPerInterval, plot(interval, steps, type="l", xlab="Interval", ylab="Steps"))
```

### Interval with most steps:
```{r biggest interval}
maxRow <- avgPerInterval[which.max(avgPerInterval$steps), ]

```
Greatest steps (`r maxRow$steps`) at interval `r maxRow$interval`.

## Imputing missing values
### Find number of rows with NAs
```{r missing values}
NROW(data) - NROW(completeData)
```

### Fill in missing values with median of dataset. Note: only "steps" column has missing values.
```{r fill}
filledData <- data
filledData$steps[is.na(filledData$steps)] <- as.integer(mean(filledData$steps, na.rm=T))
```

```{r steps per day}
filledDailySteps <- group_by(filledData, date) %>% summarize(steps=sum(steps))
qplot(filledDailySteps$steps, xlab="Number of Steps", ylab="Frequency", binwidth=1000)
```

```{r filled mean and median}
mean <- mean(filledDailySteps$steps, na.rm=T)
median <- median(filledDailySteps$steps, na.rm=T)
```
The average number of daily (with filled NAs) steps is `r mean` and the median is `r median`. 

### What is the impact of imputing missing data on the estimates of the total daily number of steps?
The mean and median went slightly down compared to the set with NAs. This implies the data is slight top heavy.

## Are there differences in activity patterns between weekdays and weekends?
```{r weekdays and weekends}
completeData$date <- as.Date(completeData$date, "%Y-%m-%d")
library(timeDate)
completeData$weekday <- isWeekday(completeData$date)
weekdaySteps <- data

avgIntervalWeekday <- filter(completeData, weekday == T) %>%
    group_by(interval) %>% 
    summarize(steps=mean(steps))
avgIntervalWeekend <- filter(completeData, weekday == F) %>%
    group_by(interval) %>% 
    summarize(steps=mean(steps))
par(mfrow=c(2,1))
with(avgIntervalWeekday, plot(interval, steps, type="l", xlab="Interbal", ylab="Steps"))
with(avgIntervalWeekend, plot(interval, steps, type="l", xlab="Interval", ylab="Steps"))

```

